Regressions And Stuff

### 2016-09-30 paper2/paper.tex
2.3

\begin{figure}[ht!]
  \centering
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/quadratic_steps_vs_mag}
    % \caption{As the step size increases, the algorithm converges quicker.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/lsq_error_gradient}
    % \caption{As $\lambda$ increases and $\delta$ decreases, error decreases.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/guess_vs_mags_2}
    % \caption{As $\mathbf{x}_0$ approaches $(10, 10)$, the rate of convergence increases.}
  \end{subfigure}
\end{figure}

@@

\begin{figure}[ht!]
  \center
  \begin{subfigure}
    \centering
      \includegraphics[width=0.45\textwidth]{../images/stochastic_gradient_steps}
    % \caption{Compare to batch gradient descent, the magnitude of the gradient during stochastic gradient descent shows significant noise.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.45\textwidth]{../images/batch_gradient_steps}
    % \caption{The magnitude of the gradient during batch gradient descent is much smoother.}
  \end{subfigure}
\end{figure}

@@

Use batch gradient descent on the SSE function on some values of M. Describe your experience
with initial guesses, step sizes and convergence thresholds. Compare with using SGD on the
same data. Explain your results in terms of the properties of the function being optimized and
the properties of the algorithms.

>>>>

\begin{figure}[ht!]
  \centering
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/quadratic_steps_vs_mag}
    % \caption{As the step size increases, the algorithm converges quicker.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/guess_vs_mags_2}
    % \caption{As $\mathbf{x}_0$ approaches $(10, 10)$, the rate of convergence increases.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.3\textwidth]{../images/lsq_error_gradient}
    % \caption{As $\lambda$ increases and $\delta$ decreases, error decreases.}
  \end{subfigure}
  \caption{The effects of $\lambda$, $\mathbf{x}_0$, and $\lambda$ on gradient descent.}
\end{figure}

@@

\begin{figure}[ht!]
  \center
  \begin{subfigure}
    \centering
      \includegraphics[width=0.45\textwidth]{../images/stochastic_gradient_steps}
    % \caption{Compare to batch gradient descent, the magnitude of the gradient during stochastic gradient descent shows significant noise.}
  \end{subfigure}
  \begin{subfigure}
    \centering
      \includegraphics[width=0.45\textwidth]{../images/batch_gradient_steps_2}
    % \caption{The magnitude of the gradient during batch gradient descent is much smoother.}
  \end{subfigure}
  \caption{Stochastic gradient descent (left) is much noisier than batch gradient descent (right).}
\end{figure}

@@

In this section, we compare the performance of batch and stochastic gradient descent on linear regression. See Figure \ref{fig:gd_m} for the performance of the two gradient descent algorithms on a polynomial function. It appears that, even for high values of $M$ (the degree of the polynomial), the results of either gradient descent are very inaccurate -- in fact, they essentially look quadratic. This is because the values of $x^k$ have extremely small values as $k$ increases for $|x| < 1$. Therefore, the gradient has a very small effect for higher degrees, while smaller degrees (i.e. $k = 0, 1, 2$) have a much larger impact.

To counteract this effect, the learning rate for the stochastic gradient descent needs to be as low as possible for. In this case, we chose $\nu_t = (r + 0.1)^{-0.51}$; with parameters any larger, the result has very little correlation with the actual function. This is likely because a large learning rate would cause smaller-degree terms to have an relatively large weight during the first few iterations, which would overtake the higher-degree terms. Having a smaller, slower-growing learning rate lets the higher-degree terms affect the result for more iterations, and therefore the curve can fit a larger number of points.

Finally, initial guess and threshold have very little impact on the accuracy of the function; they simply have an affect on performance.

\begin{figure}[ht!]
  \centering
  \label{fig:gd_m}
  \includegraphics[width=0.24\textwidth]{../images/all_m_3}
  \includegraphics[width=0.24\textwidth]{../images/all_m_5}
  \includegraphics[width=0.24\textwidth]{../images/all_m_7}
  \includegraphics[width=0.24\textwidth]{../images/all_m_10}
  \caption{The results for GD and SGD for $M = 3, 5, 7, 10$ (in that order).}
\end{figure}
